Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-07-23 concepts.tex
typos

We also consider, in measuring how much iteresting a region is, the
``innovation'' that it adds. If, during a design space exploration,
a Pareto front is temporarily calculated, adding a new Pareto front
point near to the previous ones is not a considerable ``innovation'',
because the way it fulfils the objectives is very similar to other
already evaluated configurations. On the contrary, adding a new Pareto
front point that is very distant from previous ones is remarkable,
because it let the experimenter realize that the objectives can be
fulfilled in a completely different way from what he already found
during exploration. So a region will be evaluated ineresting as much
as how different are its added Pareto front points from the ones already
found.

@@

Such an $R$ will be indicated with the following notation:
\[
R=R_{1}+R_{2}
\]
\end{defn}
\begin{rem}
$R_{1}+R_{2}$ is mathematically equivalent to $R_{1}\cup R_{2}$.
However, using $R_{1}+R_{2}$ notation, we enforce the argument that
$R_{1}$ and $R_{2}$ must be to contiuguous regions.
\end{rem}

@@

Notice that the separation is not a commutative operator.\end{defn}
\begin{rem}
The separation between $\mathbf{o}$ and $\mathbf{q}$ measures how
much we must vary $\mathbf{o}$ to transfrom it in $\mathbf{q}$.
The separation is a ``normalized'' measure that is independent of
the nature and the unit of measurement of the objectives
\end{rem}

>>>>

We also consider, in measuring how much interesting a region is, the
``innovation'' that it adds. If, during a design space exploration,
a Pareto front is temporarily calculated, adding a new Pareto front
point near to the previous ones is not a considerable ``innovation'',
because the way it fulfils the objectives is very similar to other
already evaluated configurations. On the contrary, adding a new Pareto
front point that is very distant from previous ones is remarkable,
because it let the experimenter realize that the objectives can be
fulfilled in a completely different way from what he already found
during exploration. So a region will be evaluated interesting as much
as how different are its added Pareto front points from the ones already
found.

@@

Such an $R$ will be indicated with the following notation:
\[
R=R_{1}+R_{2}
\]
\end{defn}
\begin{rem}
$R_{1}+R_{2}$ is mathematically equivalent to $R_{1}\cup R_{2}$.
However, using $R_{1}+R_{2}$ notation, we enforce the argument that
$R_{1}$ and $R_{2}$ must be to contiguous regions.
\end{rem}

@@

Notice that the separation is not a commutative operator.\end{defn}
\begin{rem}
The separation between $\mathbf{o}$ and $\mathbf{q}$ measures how
much we must vary $\mathbf{o}$ to transform it in $\mathbf{q}$.
The separation is a ``normalized'' measure that is independent of
the nature and the unit of measurement of the objectives
\end{rem}

### 2014-07-23 experiments.tex
typos

\section{A Case Study: VLIW Hardware/Software Design Space}
\label{sec:ee}
\subsection{Motivation}
In order to evaluate the behavior of the proposed approach
we will be using a parameterized Very Long Instruction
Word (VLIW) architecture~\cite{kathail_tr00} as testbed. The same philosophy
behind a VLIW system makes this a natural choice for several reasons:
\begin{itemize}
\item{Multi-Objective trade-offs}: VLIW architectures let designers
easily trade between power, energy and performances, resulting
in extended Pareto sets which are an ideal ground for our testing
purposes focused on parameter-space space representation
\item{Compiler awareness}: Instruction Level Parallelism (ILP)
is statically obtained by the compiler which
performs code trasformations being aware of the underlying hardware
configuration.  This tight hardware/software dependence, makes the
VLIW scenario perfectly suitable for investigating hardly predictable
hardware/software interactions.
\end{itemize}

@@

The results presented in the next section compare PS against one
widely adopted Multi-objective Genetic
Algorithm~\cite{knowles_techrep06}. At best of our knowledge, genetic
approaches in general still represent the best compromise between
efficiency (i.e. time required for exploration) and accuracy of the
reported Paret Sets.  Note that for the reasons described above,
mono-objective approaches have been explicitly excluded from this
analysis. Same decision has been taken for other approaches (e.g.
dependecy based~\cite{givargis_tvlsi02}) that make use of some
a-priori knowledge about the role and meaning of each parameter, since
our aim is evaluate the capacity of focusing on interesting parameter
regions without any external judgement or help based on euristics.
Finally, it should be pointed out that in this introductory work we
will explicitly use the SPEA~\cite{zitzler_eurogen01} variant of MOGA,
remainding to works like~\cite{zitzler_ec00}~\cite{zitzler_tec03} for
a detailed analysis on how it differs and what could be expected from
similar multi-objective strategies.

@@

Set of Figures~\ref{fig:pareto_fronts_100}
and~\ref{fig:pareto_fronts_500} show the resulting pareto fronts for two
different scenarios of 100 and 500 simulations budget. When possible,
similar algorithm parameters have been adopted. In particular a
population size of 30 has been chosen for GA, corresponding to the
same value for the $K$ of PS. The number of generations (or era)
has been chosen so that the only the given amount of simulations is
actually performed. Thus, when accounting these simulations only unique simulations
has been considered, removing repeations and unfeasibile
configurations.

@@

Low budget Pareto sets are shown in
Figure~\ref{fig:pareto_fronts_100}. With a so small number of
configuration being investigated the two sets are often overlapping,
being also strongly depended upon the application being considered.
These results are obtained in about 5 eras, and the main aim was
demonstrating that with a so limited budget PS still can competer with
genetic approach while even discovering very different Pareto fronts.
Continuing with a more extendent simulation budget as depicted in
Figure~\ref{fig:pareto_fronts_500}, we can qualitatively observe that
even if PS does not stricly outperforms genetic in terms of dominance, it shows in
some cases largely Pareto in terms of extension. We can intuitively
relate this to the ``novelty-based'' score system, so that instead of
focusing on getting better points inside a given range it tries to
enlarge the range itself, resulting in less dense but more extended
Paretos. Of course we cannot classify this different behavior as
``better'' or ``worse'' as compared to the genetic one, it depends on
various design factors such as desired results granularity, prefixed objective
range constraints, estimated error of the models used etc~\ldots.

>>>>

\section{A Case Study: VLIW Hardware/Software Design Space}
\label{sec:ee}
\subsection{Motivation}
In order to evaluate the behavior of the proposed approach
we will be using a parameterized Very Long Instruction
Word (VLIW) architecture~\cite{kathail_tr00} as testbed. The same philosophy
behind a VLIW system makes this a natural choice for several reasons:
\begin{itemize}
\item{Multi-Objective trade-offs}: VLIW architectures let designers
easily trade between power, energy and performances, resulting
in extended Pareto sets which are an ideal ground for our testing
purposes focused on parameter-space space representation
\item{Compiler awareness}: Instruction Level Parallelism (ILP)
is statically obtained by the compiler which
performs code transformations being aware of the underlying hardware
configuration.  This tight hardware/software dependence, makes the
VLIW scenario perfectly suitable for investigating hardly predictable
hardware/software interactions.
\end{itemize}

@@

The results presented in the next section compare PS against one
widely adopted Multi-objective Genetic
Algorithm~\cite{knowles_techrep06}. At best of our knowledge, genetic
approaches in general still represent the best compromise between
efficiency (i.e. time required for exploration) and accuracy of the
reported Paret Sets.  Note that for the reasons described above,
mono-objective approaches have been explicitly excluded from this
analysis. Same decision has been taken for other approaches (e.g.
dependency based~\cite{givargis_tvlsi02}) that make use of some
a-priori knowledge about the role and meaning of each parameter, since
our aim is evaluate the capacity of focusing on interesting parameter
regions without any external judgement or help based on heuristics.
Finally, it should be pointed out that in this introductory work we
will explicitly use the SPEA~\cite{zitzler_eurogen01} variant of MOGA,
remanding to works like~\cite{zitzler_ec00}~\cite{zitzler_tec03} for
a detailed analysis on how it differs and what could be expected from
similar multi-objective strategies.

@@

Set of Figures~\ref{fig:pareto_fronts_100}
and~\ref{fig:pareto_fronts_500} show the resulting pareto fronts for two
different scenarios of 100 and 500 simulations budget. When possible,
similar algorithm parameters have been adopted. In particular a
population size of 30 has been chosen for GA, corresponding to the
same value for the $K$ of PS. The number of generations (or era)
has been chosen so that the only the given amount of simulations is
actually performed. Thus, when accounting these simulations only unique simulations
has been considered, removing repetitions and unfeasible
configurations.

@@

Low budget Pareto sets are shown in
Figure~\ref{fig:pareto_fronts_100}. With a so small number of
configuration being investigated the two sets are often overlapping,
being also strongly depended upon the application being considered.
These results are obtained in about 5 eras, and the main aim was
demonstrating that with a so limited budget PS still can compete with
genetic approach while even discovering very different Pareto fronts.
Continuing with a more extended simulation budget as depicted in
Figure~\ref{fig:pareto_fronts_500}, we can qualitatively observe that
even if PS does not strictly outperforms genetic in terms of dominance, it shows in
some cases largely Pareto in terms of extension. We can intuitively
relate this to the ``novelty-based'' score system, so that instead of
focusing on getting better points inside a given range it tries to
enlarge the range itself, resulting in less dense but more extended
Paretos. Of course we cannot classify this different behaviour as
``better'' or ``worse'' as compared to the genetic one, it depends on
various design factors such as desired results granularity, prefixed objective
range constraints, estimated error of the models used etc~\ldots.

### 2014-07-23 introduction.tex
typos

The design of an embedded system requires different conflicting
objectives (energy consumption, performance, area) to be fulfilled.
Several factors are involved in this goal and real cases show how
often no analytical results are known to predict the relation between
system configurations and objectives. In such cases a high level
system simulation is often the only way to have a picture of how system
parameters impact on the objectives.  A key problem is that the size
of the design space to be simulated grows with the product of
cardinalities of each parameter, easily resulting in an untractable number
of simulations. In these cases a ``Design Space
Exploration'' strategy is required, i.e. a methodology to evaluate
only a small subset of all possible configurations while maintaining a
good level of efficiency and accuracy~\cite{surviving_soc}.

@@

\section{Related work}
\comment{Preferisci che i related work siano in un'unica sezione con Intro?}
Many different design space exploration algorithms have been proposed
in literature.  The motivations of an exploration
algorithm are rather heuristic, have some form of arbitrariness and,
to a large extend, intuition lies behind them.

Some exploration strategies assume some kind of knowledge about
the system parameters, their meaning and their impact on design objectives.
The \emph{Dependency analysis}, proposed in \cite{givargis_tvlsi02}, is meant
to take advantage of the parameter dependencies, a-priori known by the designer \comment{E' giusto, cosi'?}. Starting from them, he can construct a
``dependency graph'' and recognize clusters, i.e. subsets of strongly
dependency-connected parameters. Each cluster is exhaustively evaluated
and its ``local'' Pareto set is found. Then, all local Pareto sets
are merged. In this way, a series of local exhaustive evaluations
are performed instead of an exhaustive evaluation of all the possible
configurations. Some problems arise: i) In real scenarios, it may
be difficult to recognize really independent clusters of parameters,
because there may be interdependencies among a large number of
parameters. This may lead to the need of an evaluation of
almost all the possible configurations. ii) A designer may not have
a precise and complete a-priori picture of the dependencies among parameters;
for this reason the need of ``automated approaches for computing interdependencies'' is declared in the same paper. 
These drawbacks are not present in our algorithm, since, although it leverages dependencies as in \cite{givargis_tvlsi02}, it is able to detect them automatically, with no a-priori knowledge required. Moreover, our algorithm is able to capture also ``local dependencies", i.e. dependencies that
emerge only among certain ranges of parameters and may not hold when considering the entires ranges. This cannot be modeled in Dependency Analysis.

>>>>

The design of an embedded system requires different conflicting
objectives (energy consumption, performance, area) to be fulfilled.
Several factors are involved in this goal and real cases show how
often no analytical results are known to predict the relation between
system configurations and objectives. In such cases a high level
system simulation is often the only way to have a picture of how system
parameters impact on the objectives.  A key problem is that the size
of the design space to be simulated grows with the product of
cardinalities of each parameter, easily resulting in an intractable number
of simulations. In these cases a ``Design Space
Exploration'' strategy is required, i.e. a methodology to evaluate
only a small subset of all possible configurations while maintaining a
good level of efficiency and accuracy~\cite{surviving_soc}.

@@

\section{Related work and Contribution}
\label{sec:Related-work}
\comment{Preferisci che i related work siano in un'unica sezione con Intro?}
Many different design space exploration algorithms have been proposed
in literature.  The motivations of an exploration
algorithm are rather heuristic, have some form of arbitrariness and,
to a large extend, intuition lies behind them.

Some exploration strategies assume some kind of knowledge about the
system parameters, their meaning and their impact on design
objectives.  The \emph{Dependency analysis}, proposed in
\cite{givargis_tvlsi02}, is meant to take advantage of the parameter
dependencies. Starting from them, he can construct a ``dependency graph''
and recognize clusters, i.e. subsets of strongly dependency-connected
parameters. Each cluster is exhaustively evaluated and its ``local''
Pareto set is found. Then, all local Pareto sets are merged. In this
way, a series of local exhaustive evaluations are performed instead of
an exhaustive evaluation of all the possible configurations. Some
problems arise: i) In real scenarios, it may be difficult to recognize
really independent clusters of parameters, because there may be
interdependencies among a large number of parameters. This may lead to
the need of an evaluation of almost all the possible configurations.
ii) A designer may not have a precise and complete a-priori picture of
the dependencies among parameters; for this reason the need of
``automated approaches for computing interdependencies'' is declared
in the same paper.  These drawbacks are not present in our algorithm,
since, although it leverages dependencies as in
\cite{givargis_tvlsi02}, it is able to detect them automatically, with
no a-priori knowledge required. Moreover, our algorithm is able to
capture also ``local dependencies", i.e. dependencies that emerge only
among certain ranges of parameters and may not hold when considering
the entires ranges. This cannot be modelled in Dependency Analysis.
